---
title: "Vignette_Capstone"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette_Capstone}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Description

This vignette procides an overview of the R Capstone package created for visualising NOAA earthquake data, in two ways: as a timeline using a new geom derived from ggplot2 package, and as a map using leaflet package.


## Installation

To install the package, type the following:


```{r setup, include = FALSE, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r library, message=FALSE, warning=FALSE}
library(devtools)
devtools::load_all()


library(Capstone)
library(dplyr)
library(ggplot2)
library(grid)
library(lubridate)
library(leaflet)

```

## Package functions

There are 5 functions in Capstone accessible for use: 

1. eq_clean_data()
2. geom_timeline()
3. geom_timeline_label()
4. eq_map()
5. eq_create_label()

Each of the above is described below with examples. The example NOAA data can be found in the package directory under "\data" 

### eq_clean_data()

Function to clean the NOAA data, specifically uniting the YEAR, MONTH and DAY columns into a Date format, converting the LATITUDE and LONGITUDE columns from character to double numeric format and stripping out the country name from the LOCATION_NAME column. The cleaned NOAA data is returned as a data frame.

```{r Function1, message=FALSE, warning=FALSE}

df <- Capstone::noaa.data %>% Capstone::eq_clean_data()

df %>%
  dplyr::select(date, EQ_MAG_MW, DEATHS, COUNTRY, LOCATION_NAME) %>% 
  dplyr::filter(!is.na(EQ_MAG_MW) & !is.na(DEATHS)) %>%
  head(n = 10) %>% 
  print()

```

### geom_timeline()

Creates and returns a new geom to graphically depict of timelines of earthquake events. The basic geom_timeline() geom requires clean data from the previous paragraph. The required aesthetics is x with dates, optional are y for grouping by country, and size and color that can be use according to user needs. 

```{r Function2, message=FALSE, warning=FALSE}

xmindate <- ymd("2000-01-01")
xmaxdate <- ymd("2010-01-01")

df %>%
  dplyr::filter(COUNTRY %in% c("MEXICO", "INDONESIA", "CHINA")) %>%
  ggplot(aes(x = date, y = COUNTRY, label = LOCATION_NAME,group = COUNTRY, size = EQ_MAG_MW, fill = DEATHS/1000)) +
  geom_timeline(xmin = xmindate, xmax = xmaxdate) +
  # geom_timeline_label(n_max = 10, xmin = xmindate) +
  theme_classic()+ labs(y = "") +
  theme(legend.position = "bottom", axis.line.y = element_blank(), axis.ticks.y= element_blank())

```


### geom_timeline_label()

Creates and returns a new geom to label or annotate a timeline of earthquakes. This function is used in conjunction with geom_timeline() and uses an additional aesthetic "label" for labelling the earthquakes. 

```{r Function3, message=FALSE, warning=FALSE}

xmindate <- ymd("1990-01-01")
xmaxdate <- ymd("2020-01-01")

df %>%
  dplyr::filter(COUNTRY %in% c("MEXICO", "USA")) %>%
  ggplot(aes(x = date, y = COUNTRY, label = LOCATION_NAME,group = COUNTRY, size = EQ_MAG_MW, fill = DEATHS/1000)) +
  geom_timeline(xmin = xmindate, xmax = xmaxdate) +
  geom_timeline_label(n_max = 10, xmin = xmindate) +
  theme_classic()+ labs(y = "") +
  theme(legend.position = "bottom", axis.line.y = element_blank(), axis.ticks.y= element_blank())


```

### eq_map()

Creates a geographical map using OpenStreetMap leaflet tiles to display earthquake locations based on cleaned NOAA dataset

```{r Function4, message = FALSE, warning = FALSE}

df %>%
  dplyr::select(date, EQ_MAG_MW, COUNTRY, LATITUDE, LONGITUDE, TOTAL_DEATHS, LOCATION_NAME, EQ_PRIMARY) %>%
  dplyr::filter(!is.na(EQ_PRIMARY) & !is.na(TOTAL_DEATHS)) %>%
  dplyr::filter(COUNTRY %in% c("MEXICO", "USA", "CANADA")) %>%
  eq_map(data = ., annot_col = "LOCATION_NAME")

```


### eq_create_label()

Creates a html label for leaflet map based on location name, magnitude and casualties from NOAA earthquake data. The user can click on individual points on the interactive map to get details o fthe earthquake.

```{r Function5, warning=FALSE, message=FALSE}

popup_text <- eq_create_label(df)
df %>% dplyr::filter(COUNTRY %in% c("INDONESIA", "CHINA") & lubridate::year(date) >= 2000) %>%
  dplyr::mutate(popup_text = eq_create_label(.)) %>%
  eq_map(annot_col = "popup_text")

```


